```{r, echo=FALSE, results='asis'}
cat(cnt$out(ps_suffix = "Variance Components"), "\n")
```
We are given the following dataset for the trait live weight (`LiveWeight`) for cattle.

\textit{Der folgende Datensatz umfasst das Merkmal Lebendgewicht (`LiveWeight`) von Rindern.}

```{r data-table-anova, echo=FALSE, results='asis'}
knitr::kable(tbl_data_anova,
             booktabs = TRUE,
             escape = FALSE,
             format = 'latex')
```

\begin{enumerate}
\item[a)] Compute the estimate of the error variance $\sigma_e^2$ from the residuals of the fixed linear model specified below.

\textit{Schätzen Sie die Fehlervarianz $\sigma_e^2$ basierend auf den Residuen des folgenden fixen Modells.}
\points{`r lPointsQ5$TaskA`}
\end{enumerate}


The fixed linear model that is used is 

$$y = Xb + e$$
where $y$ is the vector of all live weight values, $b$ is the vector of the effects caused by the different farms. The fixed linear model is specified in R by

```{r, echo=TRUE, results='markup'}
tbl_data_anova$Farm <- as.factor(tbl_data_anova$Farm)
lm_lweight <- lm(LiveWeight ~ 0 + Farm, data = tbl_data_anova)
```

The resulting effects of the farms are 

```{r, echo=TRUE, results='markup'}
(vec_coef_lweight <- coefficients(lm_lweight))
```


\solstart

The esimate of the error variance is computed based on the resiudals. The residuals can be obtained by the function `residuals()` in R.

```{r sol-anova-a, echo=TRUE, results='markup'}
vec_res <- residuals(lm_lweight)
ssq_res <- crossprod(vec_res)
(n_est_res_var <- ssq_res / (nrow(tbl_data_anova)-length(vec_coef_lweight)))
```

The error standard deviation is 

```{r, echo=TRUE, results='markup'}
(n_est_res_sd <- sqrt(n_est_res_var))
```

\solend


\clearpage
\pagebreak



\begin{enumerate}
\item[b)] Verify your result from task a) with the output of the `summary()`-function applied to the result of the `lm()`-function

\textit{Verifizieren Sie das Resultat aus Aufgabe a) anhand des Outputs der `summary`-Funktion angewendet auf das Resultat der `lm()`-Funktion}
\points{`r lPointsQ5$TaskB`}
\end{enumerate}


\solstart

From the task, we have the result object of the `lm()`-function which is called `lm_lweight`. Applying the `summary()`-method leads to

```{r anova-sol-b, echo=TRUE, results='markup'}
summary(lm_lweight)
```

The number next to `Residual standard error: ` corresponds to the estimated value of the error standard deviation.

\solend

