---
output:
  pdf_document:
    includes:
      in_header: tex/preamble.tex
    fig_caption: false
    keep_tex: true
documentclass: article
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
library(dplyr)
```


```{r TaskCounts, echo=FALSE}
cnt <- rmdhelp::R6ClassCount$new()
cnt$set_prefix(ps_prefix = "## Problem")
```

```{r PointAssignment, echo=FALSE}
# Assign Points for Q1
lPointsQ1 <- list(TaskA = 20,
                  TaskB = 20,
                  TaskC = 5,
                  TaskD = 0)
nPointQ1Total <- sum(unlist(lPointsQ1))
# Assign Points for Q2
lPointsQ2 <- list(TaskA = 10,
                  TaskB = 30,
                  TaskC = 0)
nPointQ2Total <- sum(unlist(lPointsQ2))
# Assign Points for Q3
lPointsQ3 <- list(TaskA = 12,
                  TaskB = 6,
                  TaskC = 0)
nPointQ3Total <- sum(unlist(lPointsQ3))
# Assign Points for Q4
lPointsQ4 <- list(TaskA = 4,
                  TaskB = 10,
                  TaskC = 2)
nPointQ4Total <- sum(unlist(lPointsQ4))
# Assign Points for Q5
lPointsQ5 <- list(TaskA = 5,
                  TaskB = 5,
                  TaskC = 0)
nPointQ5Total <- sum(unlist(lPointsQ5))
# compute overal sum of points
nPointOverallTotal <- nPointQ1Total + nPointQ2Total + nPointQ3Total + nPointQ4Total + nPointQ5Total
```


\thispagestyle{empty}

\begin{tabular}{l}
ETH Zurich \\
D-USYS\\
Institute of Agricultural Sciences\\
\end{tabular}

\vspace{15ex}
\begin{center}
\huge
Solutions To Exam\\ \vspace{1ex}
Livestock Breeding and Genomics \\  \vspace{1ex}
FS 2020 \\

\normalsize
\vspace{7ex}
Peter von Rohr 
\end{center}

\vspace{7ex}
\begin{tabular}{p{5cm}lr}
  & \textsc{Date}  & \textsc{\emph{18. December 2020}} \\
  & \textsc{Begin} & \textsc{\emph{09:15 }}\\
  & \textsc{End}   & \textsc{\emph{11:15 }}\\ 
\end{tabular}

\vspace{5ex}
<!-- Table with Name -->
\large
\begin{tabular}{p{2.5cm}p{3cm}p{6cm}}
  &  Name:     &  \\
  &            &  \\
  &  Legi-Nr:  & \\
\end{tabular}
\normalsize

<!-- Table with Points -->

\vspace{9ex}
\begin{center}
\begin{tabular}{|p{3cm}|c|c|}
\hline
Problem  &  Maximum Number of Points  &  Number of Points Reached \\
\hline
1        &  `r nPointQ1Total`         & \\
\hline
2        &  `r nPointQ2Total`         & \\
\hline
3        &  `r nPointQ3Total`         & \\
\hline
4        &  `r nPointQ4Total`          & \\
\hline
5        &  `r nPointQ5Total`          & \\
\hline
Total    &  `r nPointOverallTotal`    & \\
\hline
\end{tabular}
\end{center}
