```{r, echo=FALSE, results='asis'}
cat(cnt$out(ps_suffix = "Numerator Relationship Matrix and Inbreeding"), "\n")
```

We are given the following pedigree. 

\textit{Das folgende Pedigree ist gegeben.}

\begin{center}
```{r, echo=FALSE, results='asis'}
knitr::kable(tbl_ped_nrm,
             booktabs = TRUE,
             escape = FALSE,
             format = 'latex')
```
\end{center}

\begin{enumerate}
\item[a)] Compute the additive genetic relationship matrix $A$ for the above pedigree.

\textit{Berechnen Sie die additiv-genetische Verwandtschaftsmatrix $A$ für das oben angegebene Pedigree}
\points{`r lPointsQ1$TaskA`}
\end{enumerate}

\solstart

The numerator relationship matrix is computed using `pedigreemm::getA()`. In a first step, we have to extend the pedigree to contain the founder animals in the ID-column. 

```{r sol-nrm-a, echo=TRUE, results='markup'}
vec_founder_sire <- setdiff(tbl_ped_nrm$Sire, tbl_ped_nrm$ID)
n_nr_founder_sire <- length(vec_founder_sire)
vec_founder_dam <- setdiff(tbl_ped_nrm$Dam, tbl_ped_nrm$ID)
n_nr_founder_dam <- length(vec_founder_dam)

# check that founder_sire and founder_dam are not the same
if (length(intersect(vec_founder_sire, vec_founder_dam)) != 0) 
  stop(" * ERROR: Founder sires and founder dams are not exclusive")
tbl_ped_nrm_ext <- dplyr::bind_rows(
  tibble::tibble(ID = vec_founder_sire[order(vec_founder_sire)],
                 Sex = rep('M', n_nr_founder_sire),
                 Sire = rep(NA, n_nr_founder_sire),
                 Dam  = rep(NA, n_nr_founder_sire)),
  tibble::tibble(ID = vec_founder_dam[order(vec_founder_dam)],
                 Sex = rep('F', n_nr_founder_dam),
                 Sire = rep(NA, n_nr_founder_dam),
                 Dam  = rep(NA, n_nr_founder_dam)),
  tbl_ped_nrm)

ped_nrm <- pedigreemm::pedigree(sire = tbl_ped_nrm_ext$Sire, 
                                dam = tbl_ped_nrm_ext$Dam, 
                                label = as.character(tbl_ped_nrm_ext$ID))
(matA_nrm <- as.matrix(pedigreemm::getA(ped = ped_nrm)))
```

\solend


\clearpage
\pagebreak


\begin{enumerate}
\item[b)] Compute the inbreeding coefficients of all animals in the given pedigree. Complete the following table and indicate which of the animals are inbred.

\textit{Berechnen Sie die Inzuchtkoeffizienten aller Tiere im gegebenen Pedigree. Vervollständigen Sie die folgende Tabelle und geben an, welche Tiere ingezüchtet sind.}
\points{`r lPointsQ1$TaskB`}
\end{enumerate}


\vspace{3ex}
```{r table-nrm-b, echo=FALSE, results='asis'}
n_nr_animal <- max(tbl_ped_nrm$ID)
tbl_inb_empty <- tibble::tibble(ID = 1:n_nr_animal, 
                          `Inbreeding Coefficient` = rep('', n_nr_animal),
                          `Animal Inbred (TRUE/FALSE)` = rep('', n_nr_animal))

knitr::kable(tbl_inb_empty,
             booktabs = TRUE,
             escape = FALSE,
             format = 'latex')
```

\solstart

\vspace{5ex}
The numeric solution is

```{r sol-nrm-b, echo=TRUE, results='markup'}
vec_inbr_yn <- ifelse(diag(matA_nrm) > 1, "TRUE", "FALSE")
vec_inbr_coef <- diag(matA_nrm) - 1
# tibble for table
(tbl_inb_sol <- tibble::tibble(ID = tbl_ped_nrm_ext$ID, 
                          `Inbreeding Coefficient` = round(vec_inbr_coef, digits = 3),
                          `Animal Inbred (TRUE/FALSE)` = vec_inbr_yn))
```

\solend


\clearpage
\pagebreak


\begin{enumerate}
\item[c)] Assume that dam `r dam_id` and sire `r sire_id` are mated. What is the inbreeding coefficient of their offspring?   

\textit{Wir nehmen an, dass die Mutter `r dam_id` mit dem Vater `r sire_id` angepaart wird. Wie gross ist der Inzuchtkoeffizient des Nachkommens aus dieser Paarung?}
\points{`r lPointsQ1$TaskC`}
\end{enumerate}

\solstart

The inbreeding coefficient of the offspring of dam `r dam_id` and sire `r sire_id` is half of the relationship between the parents.

```{r sol-nrm-c, echo=TRUE, results='markup'}
0.5 * matA_nrm[sire_id, dam_id]
```

\solend

