---
output:
  pdf_document:
    includes:
      in_header: tex/preamble.tex
    fig_caption: false
    keep_tex: true
documentclass: article
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
library(dplyr)
```


```{r TaskCounts, echo=FALSE}
cnt <- rmdhelp::R6ClassCount$new()
cnt$set_prefix(ps_prefix = "## Problem")
```

```{r PointAssignment, echo=FALSE}
# Assign Points for Q1
lPointsQ1 <- list(TaskA = 0,
                  TaskB = 0,
                  TaskC = 0,
                  TaskD = 0)
nPointQ1Total <- sum(unlist(lPointsQ1))
# Assign Points for Q2
lPointsQ2 <- list(TaskA = 0,
                  TaskB = 0,
                  TaskC = 0)
nPointQ2Total <- sum(unlist(lPointsQ2))
# Assign Points for Q3
lPointsQ3 <- list(TaskA = 0,
                  TaskB = 0,
                  TaskC = 0)
nPointQ3Total <- sum(unlist(lPointsQ3))
# Assign Points for Q4
lPointsQ4 <- list(TaskA = 0,
                  TaskB = 0,
                  TaskC = 0)
nPointQ4Total <- sum(unlist(lPointsQ4))
# Assign Points for Q5
lPointsQ5 <- list(TaskA = 0,
                  TaskB = 0,
                  TaskC = 0)
nPointQ5Total <- sum(unlist(lPointsQ5))
# compute overal sum of points
nPointOverallTotal <- nPointQ1Total + nPointQ2Total + nPointQ3Total + nPointQ4Total + nPointQ5Total
```


\thispagestyle{empty}

\begin{tabular}{l}
ETH Zurich \\
D-USYS\\
Institute of Agricultural Sciences\\
\end{tabular}

\vspace{15ex}
\begin{center}
\huge
Solutions To Exam\\ \vspace{1ex}
Livestock Breeding and Genomics \\  \vspace{1ex}
FS 2020 \\

\normalsize
\vspace{7ex}
Peter von Rohr 
\end{center}

\vspace{7ex}
\begin{tabular}{p{5cm}lr}
  & \textsc{Date}  & \textsc{\emph{18. December 2020}} \\
  & \textsc{Begin} & \textsc{\emph{09:15 }}\\
  & \textsc{End}   & \textsc{\emph{11:15 }}\\ 
\end{tabular}

\vspace{5ex}
<!-- Table with Name -->
\large
\begin{tabular}{p{2.5cm}p{3cm}p{6cm}}
  &  Name:     &  \\
  &            &  \\
  &  Legi-Nr:  & \\
\end{tabular}
\normalsize

<!-- Table with Points -->

\vspace{9ex}
\begin{center}
\begin{tabular}{|p{3cm}|c|c|}
\hline
Problem  &  Maximum Number of Points  &  Number of Points Reached \\
\hline
1        &  `r nPointQ1Total`         & \\
\hline
2        &  `r nPointQ2Total`         & \\
\hline
3        &  `r nPointQ3Total`         & \\
\hline
4        &  `r nPointQ4Total`          & \\
\hline
5        &  `r nPointQ5Total`          & \\
\hline
Total    &  `r nPointOverallTotal`    & \\
\hline
\end{tabular}
\end{center} 

\clearpage
\pagebreak

```{r problem setup, echo=FALSE, results='hide'}
tbl_ped_nrm <- tibble::tibble(ID = c(5,6,7,8,9,10), Sex = c('M','F','M','F','M','M'), Sire = c(3,1,5,5,7,9), Dam = c(2,4,6,6,8,8))
dam_id <- tbl_ped_nrm$Dam[length(tbl_ped_nrm$Dam)]
sire_id <- tbl_ped_nrm$Sire[length(tbl_ped_nrm$Sire)] 
```

```{r, echo=FALSE, results='asis'}
cat(cnt$out(ps_suffix = "Numerator Relationship Matrix and Inbreeding"), "\n")
```

We are given the following pedigree. 

\textit{Das folgende Pedigree ist gegeben.}


```{r, echo=FALSE, results='asis'}
knitr::kable(tbl_ped_nrm)
```

\begin{enumerate}
\item[a)] Compute the additive genetic relationship matrix $A$ for the above pedigree.

\textit{Berechnen Sie die additiv-genetische Verwandtschaftsmatrix $A$ für das oben angegebene Pedigree}
\points{`r lPointsQ1$TaskA`}
\end{enumerate}

\solstart

The numerator relationship matrix is computed using `pedigreemm::getA()`. In a first step, we have to extend the pedigree to contain the founder animals in the ID-column. 

```{r sol-nrm-a, echo=TRUE, results='markup'}
vec_founder_sire <- setdiff(tbl_ped_nrm$Sire, tbl_ped_nrm$ID)
n_nr_founder_sire <- length(vec_founder_sire)
vec_founder_dam <- setdiff(tbl_ped_nrm$Dam, tbl_ped_nrm$ID)
n_nr_founder_dam <- length(vec_founder_dam)

# check that founder_sire and founder_dam are not the same
if (length(intersect(vec_founder_sire, vec_founder_dam)) != 0) 
  stop(" * ERROR: Founder sires and founder dams are not exclusive")
tbl_ped_nrm_ext <- dplyr::bind_rows(
  tibble::tibble(ID = vec_founder_sire[order(vec_founder_sire)],
                 Sex = rep('M', n_nr_founder_sire),
                 Sire = rep(NA, n_nr_founder_sire),
                 Dam  = rep(NA, n_nr_founder_sire)),
  tibble::tibble(ID = vec_founder_dam[order(vec_founder_dam)],
                 Sex = rep('F', n_nr_founder_dam),
                 Sire = rep(NA, n_nr_founder_dam),
                 Dam  = rep(NA, n_nr_founder_dam)),
  tbl_ped_nrm)

ped_nrm <- pedigreemm::pedigree(sire = tbl_ped_nrm_ext$Sire, 
                                dam = tbl_ped_nrm_ext$Dam, 
                                label = as.character(tbl_ped_nrm_ext$ID))
(matA_nrm <- as.matrix(pedigreemm::getA(ped = ped_nrm)))
```

\solend


\clearpage
\pagebreak


\begin{enumerate}
\item[b)] Compute the inbreeding coefficients of all animals in the given pedigree. Complete the following table and indicate which of the animals are inbred.

\textit{Berechnen Sie die Inzuchtkoeffizienten aller Tiere im gegebenen Pedigree. Vervollständigen Sie die folgende Tabelle und geben an, welche Tiere ingezuchtet sind.}
\points{`r lPointsQ1$TaskB`}
\end{enumerate}

\solstart

\vspace{3ex}
```{r table-nrm-b, echo=FALSE, results='asis'}
n_nr_animal <- length(tbl_ped_nrm_ext$ID)
tbl_inb_empty <- tibble::tibble(ID = tbl_ped_nrm_ext$ID, 
                          `Inbreeding Coefficient` = rep('', n_nr_animal),
                          `Animal Inbred (TRUE/FALSE)` = rep('', n_nr_animal))

knitr::kable(tbl_inb_empty,
             booktabs = TRUE, 
             escape = FALSE,
             format = 'latex')
```

\vspace{5ex}
The numeric solution is

```{r sol-nrm-b, echo=TRUE, results='markup'}
vec_inbr_yn <- ifelse(diag(matA_nrm) > 1, "TRUE", "FALSE")
vec_inbr_coef <- diag(matA_nrm) - 1
# tibble for table
(tbl_inb_sol <- tibble::tibble(ID = tbl_ped_nrm_ext$ID, 
                          `Inbreeding Coefficient` = round(vec_inbr_coef, digits = 3),
                          `Animal Inbred (TRUE/FALSE)` = vec_inbr_yn))
```

\solend


\clearpage
\pagebreak


\begin{enumerate}
\item[c)] Assume that dam `r dam_id` and sire `r sire_id` are mated. What is the inbreeding coefficient of their offspring?   

\textit{Wir nehmen an, dass die Mutter `r dam_id` mit dem Vater `r sire_id` angepaart wird. Wie gross ist der Inzuchtkoeffizient des Nachkommens aus dieser Paarung?}
\points{`r lPointsQ1$TaskC`}
\end{enumerate}

\solstart

The inbreeding coefficient of the offspring of dam `r dam_id` and sire `r sire_id` is half of the relationship between the parents.

```{r sol-nrm-c, echo=TRUE, results='markup'}
0.5 * matA_nrm[sire_id, dam_id]
```

\solend
 

\clearpage
\pagebreak

