---
title: "Analysis of Variance"
output: html_notebook
---

# Data
The table from the slides 

```{r}
tbl_vc <- tibble::tibble(Animal = c(4:7),
                         Sire  = c(2, 1, 3, 2),
                         WWG   = c(2.9, 4.0, 3.5, 3.5))
tbl_vc
```

# Model
Model for a single observation of animal $i$ with sire $j$

$$y_i = \mu + s_{ij} + e_i$$

in matrix-vector notation

$$y = X\beta + Zs + e$$

# Components
The vector $\beta$ of fixed effects contains just one component which is the general mean $\mu$. As a consequence, the matrix $X$ has 4 rows and 1 column.

```{r}
mat_X <- matrix(1, nrow=nrow(tbl_vc))
mat_X
```

The matrix $Z$ relates observations to sire effects. In total there are three sires and therefore matrix $Z$ has three columns.

```{r}
mat_Z <- matrix(c(0, 1, 0,
                  1, 0, 0,
                  0, 0, 1,
                  0, 1, 0), nrow = nrow(tbl_vc), byrow = TRUE)
mat_Z
```

The vector of observations $y$ is taken from the datatable

```{r}
vec_y <- tbl_vc$WWG
vec_y
```


# Estimates 
Estimatesof variance components depend on $F$, $S$ and $R$. From the slides we know that 

$$F = y^TX(X^TX)^{−1}X^Ty$$

Computation of $F$

```{r}
(ytX <- crossprod(vec_y, mat_X))
(xtx <- crossprod(mat_X))
(xtxinv <- solve(xtx))
(xty <- crossprod(mat_X, vec_y))
(comp_F <- ytX %*% xtxinv %*% xty)
```

Computation of $S$

$$S = y^TZ(Z^TZ)^{−1}Z^Ty−y^TX(X^TX)^{−1}X^Ty =  y^TZ(Z^TZ)^{−1}Z^Ty - F$$

```{r}
(ytz <- crossprod(vec_y, mat_Z))
(ztz <- crossprod(mat_Z))
(ztzinv <- solve(ztz))
(zty <- t(ytz))
(comp_S <- ytz %*% ztzinv %*% zty - comp_F)
```

Computation of $R$

$$y^Ty − y^TZ(Z^TZ)^{−1}Z^Ty - y^TX(X^TX)^{−1}X^Ty= y^Ty - S - F$$

```{r}
(yty <- crossprod(vec_y))
(comp_R <- yty - comp_S - comp_F)
```

# Results
Estimates are obtained by replacing expected values of componets with observed values and replacing variance components by their estimates (method of moments)

$$\widehat{\sigma_e^2} = \frac{R}{n-q}$$
with $n$ is the number of observations and $q$ is the number of sires.

```{r}
(est_err_var <- comp_R / (4-3))
```

The estimate for $\sigma_s^2$ is obtained from

$$M = I − X(X^TX)^{−1}X^T$$

```{r}
(xxtxinvxt <- mat_X %*% xtxinv %*% t(mat_X))
(mat_M <- diag(1,nrow = nrow(tbl_vc)) - xxtxinvxt)
(ztmz <- t(mat_Z) %*% mat_M %*% mat_Z)
trztmz <- sum(diag(ztmz))
(est_sire_var <- (comp_S - (3-1)*est_err_var) / trztmz)
```


# In R
R provides the function `aov()` which can be used to get to these results. For this the sire effect is converted into a fixed effect by changing its data type into a factor. 

```{r}
tbl_vc$Sire <- as.factor(tbl_vc$Sire)
aov_sire <- aov(WWG ~ 1 + Sire, data = tbl_vc)
summary(aov_sire)
```


# Likelihood 
Likelihood-based approach can be done with the pedigreemm package. For our example dataset

```{r}
n_nr_sire <- nlevels(as.factor(tbl_vc$Sire))
ped_sire <- pedigreemm::pedigree(sire = c(NA, NA, NA), dam = c(NA, NA, NA), label = as.character(1:n_nr_sire))
ped_sire
```

The pedigreemm package contains a function `pedigreemm` which computed predicted breeding values and variance components. 

```{r}
lme_sire <- pedigreemm::pedigreemm(formula = WWG ~ 1 + (1|Sire),
                                   pedigree = list(Sire = ped_sire), 
                                   data = tbl_vc)
summary(lme_sire)
```

