---
title:  Livestock Breeding and Genomics - Exercise 2
author: Peter von Rohr
date: 2020-10-02
output: 
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE, results = 'asis')
```


## Problem 1: Breeding Values For a Monogenic Trait
We assume that the absorption of cholesterol is determined by a certain enzyme. The level of enzyme production is determined by a single bi-allelic locus $E$. The genotype frequencies and the genotypic values for the two dairy cattle populations `Original Braunvieh` and `Brown Swiss` are given in the following table.

```{r table-geno-freq-value, echo=FALSE}
l_maf <- list(ob = 0.25, bs = 0.1)
l_a <- list(ob = 15, bs = 29)
l_d <- list(ob = 3, bs = 0)
tbl_geno_freq_value <- tibble::tibble(Variable = c(
  "$f(E_1E_1)$",
  "$f(E_1E_2)$",
  "$f(E_2E_2)$",
  "$a$",
  "$d$"),
  `Original Braunvieh` = c(
    l_maf$ob^2, 
    2*l_maf$ob*(1-l_maf$ob),
    (1-l_maf$ob)^2,
    l_a$ob,
    l_d$ob
  ),
  `Brown Swiss` = c(
    l_maf$bs^2,
    2*l_maf$bs * (1-l_maf$bs),
    (1-l_maf$bs),
    l_a$bs,
    l_d$bs
  ))
knitr::kable(tbl_geno_freq_value, booktabs = TRUE, escape = FALSE)
```


### Hints
* Assume that allele $E_1$ is the allele with the positive effect on the enzyme level
* Assume that the Hardy-Weinberg Equilibrium holds in both populations


### Your Task
Compute the breeding values for all three genotypes in both populations. 


### Your Solution

\pagebreak

## Problem 2: Matrices in R
In R, matrices are constructed using the function `matrix()`. This function accepts different options. We want to see, how these options work.

__Your Task__: Construct matrices using the different options to better understand the meaning of the different options.


### Parameter `data`
* `data`: Specify the different matrix elements
```{r echo=TRUE, results='markup', eval=FALSE}
(matA <- matrix(data = c(1:9), nrow = 3, ncol = 3))
```

* `data`: without specifying the matrix elements
```{r echo=TRUE, results='markup', eval=FALSE}
(matB <- matrix(nrow = 3, ncol = 3))
```

* `data`: specifying not all matrix elements
```{r echo=TRUE, results='markup', eval=FALSE}
(matC <- matrix(data = c(1,2,3), nrow = 3, ncol = 3))
```

```{r echo=TRUE, results='markup', eval=FALSE}
(matC2 <- matrix(data = c(1,2,3,4), nrow = 3, ncol = 3))
```


### Parameters `nrow` and `ncol`
* Leaving out one of both parameters

```{r echo=TRUE, results='markup', eval=FALSE}
(matD <- matrix(data = c(1:9), nrow = 3))
```

```{r echo=TRUE, results='markup', eval=FALSE}
(matE <- matrix(data = c(1:9), ncol = 3))
```

### Parameter `byrow`
```{r echo=TRUE, results='markup', eval=FALSE}
(matF <- matrix(data = c(1:9), nrow = 3, ncol = 3, byrow = TRUE))
```

```{r echo=TRUE, results='markup', eval=FALSE}
(matG <- matrix(data = c(1:9), nrow = 3, ncol = 3, byrow = FALSE))
```


## Problem 3: Matrix multiplication in R
In R, matrices can be multiplied using the operator `%*%` or with the functions `crossprod()` or `tcrossprod()`. With `crossprod()` and `tcrossprod()` vectors and matrices can be multiplied directly. The conversion of vectors to matrices is done automatically inside of these functions. The result will always be a matrix. When doing matrix-vector multiplications with `%*%` the vector has to be converted first into a matrix using the function `as.matrix()`.

In a first part of this problem, compare the results of the functions `crossprod()`, `tcrossprod()` and `%*%`. 

a) Given are the following matrices

```{r MatMultMatDef, echo=TRUE, results='markup'}
matA <- matrix(data = c(1:9), ncol = 3)
matB <- matrix(data = c(2:10), ncol = 3)
```

Find out which matrix multiplication with `%*%` corresponds to the following statement?

```{r, echo=TRUE, results='markup', eval=TRUE}
crossprod(matA,matB)
```

### Your Solution


b) Given is the vector `vecB`

```{r, echo=TRUE}
vecB <- c(-3,16,1)
```

Multiply the matrix `matA` with the vector `vecB` once using `%*%` and then with the function `crossprod()`.

__Hint__: a vector can be converted to a matrix using the function `as.matrix()`.


### Your Solution

## Problem 4: Quantitative Genetics
In a population the following numbers of genotypes were counted for a given genetic locus called $A$. 

```{r GenotypeFreqs}
dfGenotypeFreq <- data.frame(Genotypes = c("$A_1A_1$", "$A_1A_2$", "$A_2A_2$"),
                             Numbers   = c(24, 53, 23),
                             stringsAsFactors = FALSE)
knitr::kable(dfGenotypeFreq)
```

a) Compute the genotype frequencies

### Your Solution

b) Compute the allele frequencies

### Your Solution

c) Compute the population mean $\mu$ under the following assumptions

* the difference between the genotypic values of the homozygous genotypes is $20$ and
* the genotypic value of the heterozygous genotype is $2$.

### Your Solution

