---
title: "Example Pedigree"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Definition of pedigree in package pedigreemm

```{r}
ped <- pedigreemm::pedigree(sire = c(NA,NA,NA,1,3,4),
                           dam  = c(NA,NA,NA,2,2,5),
                           label = as.character(1:6))
ped
```

The numerator relationship matrix $A$

```{r}
(sparse_A = pedigreemm::getA(ped = ped))
```

The cholesky decomposition of $A$ 

```{r}
A <- as.matrix(sparse_A)
(R <- t(chol(A)))
```

To get to the matrix $L$, we have to use the definition of $R = L * S$ and $S$ is the diagnoal matrix with elements corresponding to the square root of matrix $D$. Matrix $D$ can be obtained as 

```{r}
(D <- diag(pedigreemm::Dmat(ped = ped), nrow = 6))
```

Matrix $S$ is obtained from matrix $D$

```{r}
(S <- sqrt(D))
```

The matrix $L$ is obtained from the defintion of matrix $R = L * S$, therefore the matrix $L = R * S^{-1}$

```{r}
(L <- R %*% solve(S))
```

The matrix $P$ that we used in the simple decomposition, is computed from the relationship $L^{-1} = I-P$ and from that we get $P = I - L^{-1}$

```{r}
(P <- round(diag(1,nrow = 6) - solve(L), digits = 2))
```

The inbreeding coefficients $F_i$ for all animals in the pedigree, can also be obtained with pedigreemm

```{r}
pedigreemm::inbreeding(ped = ped)
```

